---
title: "Experiment 3: More Airline Data"
subtitle: "Asynchronous lecture"
date: "2023-09-10"
output: html_notebook
---
The ecperiment (conceotually) continues our class work on flight data.
We will expand our dataset by going to a root source - the Bureau of Transportation

# Packages 

```{r}
library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
```

# Load BOT data

```{r}
thisfile = here("T_T100D_SEGMENT_US_CARRIER_ONLY.CSV")

df2022 = read_csv(thisfile) %>% clean_names()
```

## Subsetting to data of interest

Let's focus on flights from La Guardia (airport "LGA") and eliminate cargo flights by requiring at least 1 passenger.

Call the resultant dataframe df9.

```{r}
df9 = df2022 |> 
  filter(passengers>0, origin == "LGA")
```

# Exploring the BOT 2022 data

## Subsetting the data of interest

Lets try to visualize a monthly count of flights using a barchart.

First, a not quite right visualiztion.

```{r}
df9 |> 
  ggplot(aes(month)) + geom_bar()
```

Counting rows is not going to give us a count of flights. The cisualization we want is:

```{r}
df9 |> 
  ggplot(aes(month)) + geom_bar(aes(weight = departures_performed))
```

Observation:
* low passenger numbers in early 2022 likely due to covid.

###Just because you can

```{r}
df9 |> 
  ggplot(aes(month, fill=carrier_name)) + geom_bar(aes(weight=departures_performed))
```

## Arrivals and departures from LGA

```{r}
df10 = df2022 |> filter(passengers>0, origin=="LGA" | dest=="LGA")

df10 |> ggplot(aes(month)) + geom_bar(aes(weight = passengers))
```

```{r}
df11 = df10 |>
  select(month, passengers, seats, carrier_name, dest, origin)
```

```{r}
df12 = df10 |> select(1:5, month, contains("id"))
```

```{r}
df13 = df11 |> mutate(percent_loading = passengers/seats*100)

df13 |> ggplot(aes(percent_loading)) +
  geom_histogram() + facet_wrap(~carrier_name, scales="free_y")
```


















