---
title: "Experiment 1: Airline Data Analysis"
Date: "2023-09-06"
output: html_notebook
---

First in-class DS241 project; loosely adopted from 'R for Data Science' textbook.

```{r}

library(tidyverse)
library(janitor)
library(here)
library(nycflights13)
library(ggplot2)

# Loading raw (tidied) data
df1 = nycflights13::flights
glimpse(df1)
```

Task 1:
```{r}

df2 = df1 |> filter(month == 9, origin == "MIA")
glimpse(df2)

```

Task 2:
```{r}
df3 = df1 |> filter(month == 9, dest == "MIA")
glimpse(df3)
  
```

Task 3:
```{r}
df4 = df1 |> filter(month == 1, dest == "MIA")
glimpse(df4)
```

Task 4:
```{r}
df5 = df1 |> filter(5 < month, month < 9, dest %in% c("MDW", "ORD"))
glimpse(df5)
```

```{r}
numbers = unique(df3$flight)
smallflight = min(numbers)
```

```{r}
df6 = df1 |> filter(dest == "MIA", flight == smallflight)
```

```{r}
ggplot(data = df6, 
       mapping = aes(x = dep_delay, y = arr_delay)) + geom_point()
```

```{r}
ggplot(data = df6, 
       mapping = aes(x = dep_delay, y = arr_delay, colour = month)) + geom_point() 
```
It is noticeable in this plot comparing the arrival and departure delays that greater delays are more frequent in earlier months of the year. Most of the smallest arrival and departure delays are from September, October, and November.


Extra Credit:
"Conversely, negative delays (early arrivals) do not exhibit any power laws but simply behave in an exponential way, with extremely early arrivals exponentially unlikely." 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8041857/
